%
\section{UI framework}
\label{sec:ui-framework}

For the development of the \texttt{Local System} and the \texttt{Remote Client}, it is necessary to use a \gls{ui} Framework, in order to develop a \gls{ui}, making it more user friendly and interactive.
There are several frameworks that can be used in Linux, such as~\cite{ui-lists}:
%
\begin{item-c}
\item Qt;
\item Sciter;
\item Noesis GUI;
\item wxWidgets;
\item GTK+;
\item and so on.
\end{item-c}

\subsection{Qt}

The Qt framework contains a comprehensive set of highly intuitive and modularized C++ library classes and is loaded with APIs to simplify application development~\cite{qt-site}.

\subsubsection{Signals and slots}

In Qt, there's an alternative to the callback technique, using \texttt{signals and slots}. A \texttt{signal} is emitted when a particular event occurs and a \texttt{slot} is a function that is called in response to a particular signal~\cite{qt-signals-slots}.

\begin{itemize}
\item\emph{Signals}: Are emitted by an object when its internal state has changed in some way that might be interesting to the object's client or owner. Signals are also public access functions and can be emitted from anywhere~\cite{qt-signals-slots}.
\item\emph{Slots}: A slot is called when a signal connected to it is emitted. Slots are normal C++ functions and can be called normally; their only special feature is that signals can be connected to them~\cite{qt-signals-slots}.
\end{itemize}

Compared to callbacks, signals and slots are slightly slower because of the increased flexibility they provide, although the difference for real applications is insignificant~\cite{qt-signals-slots}.

\subsubsection{Usage Example}



For this project, \texttt{Qt} was chosen due to the following reasons:
\begin{item-c}
\item
  First of all, it is possible to `develop graphical user interfaces and cross-platform applications, both desktop and embedded'~\cite{qt-bib}.
The framework operates on different types of software and hardware.
Also, it has an easy-to-read code and it is possible to make an attractive
\gls{ui}.
\item 
Secondly, this framework is \texttt{cost-friendly}, not only because it has a
free license, but also because its software development takes less time to
develop due to its highly productive features.
\item
Lastly, it is implemented in C++, which means that it is possible to use many libraries.
The wide choice of modules allow the project to have rich functionality and as a result, the software will have a \gls{gui} similar to a native one.
\end{item-c}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../../dissertation"
%%% End:
